Lesson 5
========================================================
=)
```{r}
pf<-read.csv(file = "pseudo_facebook.tsv",sep = '\t',dec=".")
pf
```

### Multivariate Data
Notes:
3 or more variables at a time
***

### Moira Perceived Audience Size Colored by Age
Notes:
***
Scatter plot by age, it does not help
Third level addes color
There is no pattern


### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(ggplot2)
library(gridExtra)

histogram1<-ggplot(aes(x = gender, y = age),data = subset(pf, !is.na(gender))) + geom_histogram(stat = "identity")

boxplot1<-ggplot(aes(x = gender, y = age),data = subset(pf, !is.na(gender))) + geom_boxplot()+stat_summary(fun.y=mean,geom = 'point',shape=4)

grid.arrange(histogram1,boxplot1)
```

```{r}
ggplot(aes(x = age, y = friend_count),data = subset(pf, !is.na(gender))) + geom_line(aes(color=gender),stat="summary",fun.y=mean)
```
```{r}
library(dplyr)
pf.fc_by_age_gender<-pf %>%  group_by(age,gender) %>%
  filter(gender!="NA") %>% summarise(mean_friend_count=mean(friend_count),median_friend_count=median(friend_count),n=n())  %>% arrange(age)
head(pf.fc_by_age_gender)
```



***
### Plotting Conditional Summaries
Notes:
```{r Plotting Conditional Summaries}
ggplot(aes(x=age,y=mean_friend_count),data=pf.fc_by_age_gender)+ geom_line(aes(color=gender))+geom_smooth()

```

***

### Thinking in Ratios
Notes:
Looking at this plot it seems that for Young people the difference is higher for females than male
***

### Wide and Long Format
Notes:
  Another approach would be better to understand our data
  We will use Reshape2 to do that, we are agoing to convert
  gender into two columns
***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age~gender, value.var='median_friend_count')
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x=age,y=female/male),data=pf.fc_by_age_gender.wide)+geom_line()+geom_hline(yintercept=1,color="red",linetype=2,alpha=0.3) 
```
***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined<-floor(2014- pf$tenure/365)

```
***

### Cut a Variable
Notes:
```{r Cut a Variable}
pf$year_joined.bucket<-cut(pf$year_joined,c(2004,2009,2011,2012,2014))
```
***
Cut entries first we need the field we want to cut and the next input is the vector or the number of breaks that our field will be separate


### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x=age,y=friend_count),data=subset(pf,!is.na(year_joined.bucket)))+ geom_line(aes(color=year_joined.bucket),stat='summary',fun.y=median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x=age,y=friend_count),data=subset(pf,!is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket),stat='summary',fun.y=median)+
  geom_line(stat='summary',fun.y=mean,linetype=6,color="blue")
```

***

### Friending Rate
Notes:
```{r Friending Rate}
library(dplyr)
with(subset(pf,tenure>=1), summary(friend_count/tenure))
```
***

### Friendships Initiated
Notes:

What is the median friend rate?
0.2205
What is the maximum friend rate?
417.0000  <- OUTLIER  cannot be possible because our third quartile is under 0.6


```{r Friendships Initiated}
ggplot(aes(x=tenure,y=friendships_initiated/tenure),data=subset(pf,tenure >= 1))+geom_line(aes(color=year_joined.bucket),stat='summary',fun.y=mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***
```{r}
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```

### Sean's NFL Fan Sentiment Study
Notes:
***


### Introducing the Yogurt Data Set
Notes:

***
```{r}
yogurt_dts<-read.csv('yogurt.csv',sep = ",")
yogurt_dts$id<-factor(yogurt_dts$id)
str(yogurt_dts)
```

### Histograms Revisited
Notes:

```{r Histograms Revisited}
library(ggplot2)
ggplot(aes(yogurt_dts$price),data=yogurt_dts)+geom_histogram()
hist(yogurt_dts$price)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
yogurt_dts<-transform(yogurt_dts,all.purchases=strawberry+blueberry+pina.colada+plain+mixed.berry)
summary(yogurt_dts$all.purchases)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x=time,y=price),data=yogurt_dts)+geom_jitter(alpha=1/4,shape=21,fill=I('#F79420'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(12)
dtsSample<-sample(levels(yogurt_dts$id),16)

ggplot(aes(x=time,y=price),data=subset(yogurt_dts, id %in% dtsSample)) + facet_wrap(~id) +geom_line() + geom_point(aes(size=all.purchases),pch=1)


```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:
Let these chosen variables speak, sometimes these variables are 20- 100
so one way to get idea of the relations we can speed up our
exploratory analysis by producing many plots
***

### Scatterplot Matrix
Notes:
There is a grid of scatter plots between every pair of variables
***
```{r}
#install.packages("GGally")
library(GGally)
theme_set(theme_minimal(6))
set.seed(1836)
pf_subset<-pf[,c('age','friend_count','friendships_initiated','mobile_likes')]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])

```


### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

